---
title: The tomato experience
knitr:
  opts_chunk: 
    echo: false
    message: false
    warning: false
    comment: NA
    cache: false
    fig.width: 7
    fig.asp: 0.5
    fig.path: "figures_tomatoes/"
    dpi: 320
    out.width: 100%
    external: false
    res: 320
    results: "hide"
    split: false 
    tidy: false
    dev: svg
editor_options: 
  chunk_output_type: console
format: 
  html: 
    embed-resources: true
---


```{r setup}
library(httr)
library(apastats2)
library(circhelp)
library(data.table)
library(jsonlite)
library(stringr)
library(knitr)
library(Hmisc)
library(ez)
library(superb)

default_font = 'Gill Sans Nova Light'

default_font_size <- 10
default_line_size <- 1/.pt/3.82*4 
default_font_size_mm <- default_font_size/ggplot2:::.pt

default_point_size <- 4*default_line_size

update_geom_defaults("line", list(linewidth = default_line_size))
update_geom_defaults("text", list(size = default_font_size_mm, family = default_font))
update_geom_defaults("segment", list(linewidth = default_line_size))
update_geom_defaults("pointrange", list(linewidth = default_line_size))
update_geom_defaults("vline", list(linewidth = default_line_size, colour = '#AFABAB'))
update_geom_defaults("hline", list(linewidth = default_line_size, colour = '#AFABAB'))
update_geom_defaults('function', list(linewidth = default_line_size))

default_theme<-theme_light(base_size = default_font_size, base_line_size = default_line_size, base_family = default_font)+theme(
  axis.line=element_line(linewidth = I(0.5)), 
  axis.ticks= element_line(linewidth = I(0.25), colour = 'gray'),
  axis.line.x=element_line(),
  axis.line.y=element_line(),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  legend.title=element_text(size=rel(1)), 
  strip.text=element_text(size=rel(1), colour = 'black'), 
  axis.text=element_text(size=rel(0.9), colour = '#4e4e4e'), 
  axis.title=element_text(size=rel(1), colour = '#4e4e4e'), 
  panel.border= element_blank(),
  strip.background = element_blank(),
  legend.position	='right', 
  legend.key.height = unit(0.7, 'lines'),
  plot.title=element_text(size=rel(1), hjust = 0.5),
  plot.tag.position = c(0,1),
  plot.background = element_blank(),
  text=element_text(size=default_font_size), 
  legend.text=element_text(size=rel(1)), 
  axis.line.x.bottom = element_blank(), 
  axis.line.y.left = element_blank(),
  axis.line.x.top = element_blank(),
  axis.line.y.right = element_blank())
theme_set(default_theme)

default_colors <-c('#3498db','#009f06','#FF7F00', '#7A378B','#f72034')
scale_colour_discrete <- function(...) scale_colour_manual(values=default_colors, ...)
scale_fill_discrete <- function(...) scale_fill_manual(values=default_colors, ...)
scale_shape_discrete <- function(...) scale_shape_manual(values = c(16,21,17, 22), ...)

```

```{r}

headers = c(
  `Authorization` = sprintf("Bearer %s",Sys.getenv('JATOS_TOKEN')),
  `Content-Type` = 'text/plain'
)

jatos_url <- 'https://jatos.mindprobe.eu/jatos/api/v1/'
res <- httr::GET(url = paste0(jatos_url, "admin/token"), httr::add_headers(.headers=headers))

randstring <- function(n = 5000) {
  a <- do.call(paste0, replicate(5, sample(LETTERS, n, TRUE), FALSE))
  paste0(a, sprintf("%04d", sample(9999, n, TRUE)), sample(LETTERS, n, TRUE))
}

temp_dir <- tempdir(check = T)
temp_dir <- gsub('\\\\', '/', temp_dir)
temp_dir <- paste(temp_dir, randstring(1), sep = '/')
res_file <- paste0(tempfile(),'.zip')

study_uid <- Sys.getenv('JATOS_TOMATO_STUDY_2026_UID')

req_url <-  sprintf("%sresults/metadata?studyUuid=%s", jatos_url, study_uid)
system(sprintf('curl  -H "Authorization: %s" %s --output "%s" ', headers[['Authorization']], req_url, res_file))

metadata_global <- jsonlite::fromJSON(res_file)$data$studyResults[[1]]
setDT(metadata_global)
unlink(paste0(temp_dir, "/*"))
finished_study <- metadata_global[studyState=='FINISHED',c('id','startDate','endDate','duration')]
finished_study <- finished_study[as.POSIXct(startDate/1000, origin="1970-01-01")>Sys.getenv("JATOS_START_DATE")]
url <-  sprintf("%sresults?studyUuid=%s", jatos_url, study_uid)
system(sprintf('curl -i -H "Authorization: %s" %s --output "%s"', headers[['Authorization']], url, res_file))
unzip(res_file, exdir=temp_dir)  # unzip your file

metadata <- fromJSON(list.files(temp_dir, '*.json', full.names = T, recursive = T))$data$studyResults[[1]]

finished_study_ids <- metadata[,c('id', "endDate", "startDate")]
setDT(finished_study_ids)
finished_study_ids[,id:=as.character(id)]
finished_study_ids[,endDate:=as.POSIXct(endDate/1000, origin="1970-01-01")]
finished_study_ids[,startDate:=as.POSIXct(startDate/1000, origin="1970-01-01")]
finished_study_ids <- finished_study_ids[startDate>Sys.getenv("JATOS_START_DATE")]

comp_res_meta <- rbindlist(lapply(1:length(metadata$componentResults), \(x) {
  dt <- as.data.table(metadata$componentResults[[x]])
  dt
}), fill = T)
flist <- list.files(temp_dir, '*.txt', full.names = T, recursive = T)
flist <- flist[str_extract(flist,'study_result_(\\d+)', group = 1)%in%finished_study_ids$id]

cur_data <- rbindlist(lapply(c(flist), \(x){
  print(x)
  comp_res_id <- str_extract(x, '(?<=comp-result_)\\d{6}')
  data <- read_json(x)
  if (length(data)<40){
    return(data.frame())
  }
  browser_check <-data[sapply(data, \(x) x$trial_type=='browser-check')][[1]]

  setDT(browser_check)
  
  data <- data[sapply(data, \(x) x$trial_type=='psychophysics')]
  data <- rbindlist(data, fill = T)
  if ('score' %nin% colnames(data)){
    return(data.frame())
  }

    km <- kmeans(as.matrix(data[!is.na(click_x),.(click_x,click_y)]), 3, nstart = 50)
    sorted_indices <- order(km$centers[, 1])  # Assuming x-coordinate is in the first column
    sorted_indices[2:3] <- rev(sorted_indices[2:3])
        
    # Reassign cluster labels based on sorted order
    new_clusters <- match(km$cluster, sorted_indices)
    data[!is.na(click_x), cue_n_alt:=new_clusters]
    # data[, cue_pos:=ifelse(cue_n==1, 'left', ifelse(cue_n==2, 'right', 'avg'))]
    
  if ('cue_n' %nin% names(data)){
    data[,cue_n:=cue_n_alt]
  } 
  data[,score:=as.numeric(score)]
  data[!is.na(left_colors), left_sd:=circ_sd_360(as.numeric(str_split(left_colors,',', simplify = T))), by = left_colors]
  data[!is.na(right_colors), right_sd:=circ_sd_360(as.numeric(str_split(right_colors,',', simplify = T))), by = right_colors]
  data[,left_colors:=NULL]
  data[,right_colors:=NULL]
  setnafill(data, type = "locf",cols = c('target_color','distractor_color','left_color','right_color','left_population_mean','right_population_mean', 'left_correlation', 'right_correlation','base_ori','overall_sample_mean','similarity', 'similarity_bin', 'color_bin','left_sd','right_sd','cue_n'))
  setnafill(data, type = "nocb",cols = c('score'))
  setnames(data,c('base_ori','left_color','right_color'),c('base_color','left_sample_mean','right_sample_mean'))
  #
  if (browser_check[1, .(trial_type)]=='browser-check'){
    data<-cbind(data,browser_check[,.(width, height, fullscreen)])
  }
  data<-data[trial_type=='psychophysics'&seq_type=='response']
  # data[, error:=angle_diff_360(resp_color, target_color)]
  data[,subject_id:=str_extract(x, 'study_result_(\\d+)', group = 1)]
  data[,result_id:=str_extract(x, 'comp-result_(\\d+)', group = 1)]
  data <- merge(data, finished_study_ids, by.x = 'subject_id', by.y = 'id', all.x = T)
  
  data
  
}), fill = T)
```

```{r}
#| label: data-preprocessing

tomatoes <- cur_data
tomatoes[,expName:='tomato_experience_2026']
tomatoes[,.(left_population_mean, right_population_mean, base_color,
            target_color, distractor_color, left_sample_mean, right_sample_mean)]
tomatoes[,realErr:=angle_diff_360(resp_color, base_color )]
tomatoes[left_correlation!=right_correlation, c('error_to_uncorrelated','error_to_correlated') := .(
  angle_diff_360(resp_color, fifelse(left_correlation<right_correlation,  left_sample_mean, right_sample_mean)), 
  angle_diff_360(resp_color, fifelse(left_correlation>right_correlation,  left_sample_mean, right_sample_mean)))]

# correlation level defines variability level
# (correlated = low variability, uncorrelated = high variability).
tomatoes[,item_low_var_color:=fifelse(
  left_correlation>right_correlation, left_sample_mean,
  right_sample_mean
)]
tomatoes[,item_high_var_color:=fifelse(
  left_correlation<right_correlation, left_sample_mean,
  right_sample_mean
)]

tomatoes[,abs_err:=abs(error)]
tomatoes <- unique(tomatoes)
tomatoes[!is.na(error),n_by_subj:=.N, by = subject_id]
tomatoes[!is.na(error),n_by_block:=.N, by = .(result_id, subject_id)]

tomatoes[,color_dist:=angle_diff_360(item_low_var_color, item_high_var_color)]
tomatoes[,abs_color_dist:=abs(color_dist)]
tomatoes[,similarity_binf:=factor(similarity_bin)]

tomatoes<-tomatoes[n_by_subj>250&block>0]

aggr_stats <- tomatoes[,.(totalN = .N, non_missing=n_by_subj[1], 
                          circ_sd_err_to_pop_mean = circ_sd_360(realErr, na.rm = T), 
                          circ_sd_err_to_sample_mean = circ_sd_360(error, na.rm = T),
                          circ_sd_err_to_uncorrelated = circ_sd_360(error_to_uncorrelated, na.rm = T),
                          circ_sd_err_to_correlated = circ_sd_360(error_to_correlated, na.rm = T),
                          n_by_subj[1], missed_responses = sum(is.na(error))), keyby = .(subject_id)]
fwrite(tomatoes, file = "tomatoes_raw.csv")                
```

```{r results='hide'}
tomatoes[,.(.N, circ_sd = circ_sd_360(error, na.rm = T), mean_abs_err= mean_nn(abs_err)), keyby = .(startDate, endDate,  subject_id)]
plot(density(tomatoes[!is.na(error), error]))
ggplot(tomatoes[!is.na(error)], aes(x = error))+
  geom_density()+
  geom_rug()
  labs(x = 'Error (degrees)', y = 'Density')+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r eval = T}

#outliers 
tomatoes <- tomatoes[subject_id%nin%tomatoes[,.(.N, circ_sd_360(error, na.rm = T)), by = subject_id][V2>60|N<200, subject_id]]
tomatoes <- tomatoes[cue_n == 3]
tomatoes[,subject:=paste0('S',as.numeric(as.factor(subject_id)))]
tomatoes[,subject:=naturalsort::naturalfactor(subject)]

tomatoes[,.(.N, circ_sd_360(error, na.rm = T)), by = subject]
db_res <- lapply(split(tomatoes[!is.na(error),], by = c('subject_id')), \(dt){
  dt[,c('be_c','is_outlier','rc_pred','rc_group','shifted_color_dist'):=circhelp::remove_cardinal_biases(error, overall_sample_mean , space = '360', plots = 'hide')[,c('be_c','is_outlier','pred','which_bin','shifted_x')], by = .(subject)]
  dt
})

tomatoes <-rbindlist(db_res)
tomatoes <- tomatoes[!is.na(abs_err)&is_outlier==F]
```

```{r}
tomatoes[is_outlier==F&abs_err>90, is_outlier:=T]

tomatoes[!is.na(abs_err), 
         bias_rel_to_high_var := sign(angle_diff_360(item_high_var_color, base_color)) * be_c]

aggr_stats <- tomatoes[,.(.N, circ_sd = circ_sd_360(error, na.rm = T),circ_sd_correct = circ_sd_360(be_c, na.rm = T), n_by_subj=n_by_subj[1], n_nonmissing=sum(!is.na(error)), share_outliers= mean_nn(is_outlier), mean_abs_err= mean_nn(abs_err), bias_rel_to_high_var = circ_mean_360(bias_rel_to_high_var, na.rm = T)), keyby = .(subject)]#[,mean_nn(share_outliers)]

tomatoes[,abs_be_c:=abs(be_c)]
tomatoes[,avg_abs_err:=mean_nn(abs(be_c)), by = subject]
corr_midpoint <- mean(range(c(tomatoes$left_correlation, tomatoes$right_correlation), na.rm = TRUE))
tomatoes[,left_variability := fifelse(left_correlation <= corr_midpoint, "high", "low")]
tomatoes[,right_variability := fifelse(right_correlation <= corr_midpoint, "high", "low")]
tomatoes[,noise_equal:=factor(left_variability == right_variability, levels = c(FALSE, TRUE), labels = c('Unequal noise', 'Equal noise'))]
tomatoes[,variability_condition:=factor(paste0(left_variability, " - ", right_variability),
                                        levels = c("low - low", "high - low", "low - high", "high - high"))]
tomatoes[,similarity_binf:=factor(similarity_bin)]

tomatoes <- drop_empty_cols(tomatoes)

tomatoes[,c('startDate','endDate','result_id','subject_id','avg_frame_time'):=NULL]

if (lengthu(tomatoes$subject)>1){

tomatoes <- tomatoes[, lapply(.SD, function(col) if (!all(na.omit(col == col[!is.na(col)][1]))) col else NULL)]
}
tomatoes_out <- tomatoes[order(subject, trial_number),.(subject, trial_number, low_var_color = item_low_var_color, high_var_color = item_high_var_color, left_color = left_sample_mean, right_color = right_sample_mean, left_sd,  right_sd, variability_condition, similarity_bin, resp_color,  rawError = error, base_color , realErr, errorCorrected = be_c, bias_rel_to_high_var, is_outlier)]
fwrite(tomatoes_out, file = 'tomatoes.csv')
```

# Instructions

Read the report below and use it as a basis for your presentation. You can also download the preprocessed data from `tomatoes.csv` and use your favorite software (JASP, jamovi, R, SPSS) to analyze it to get more insight into the study results. The data file has the following columns:

* `subject` - participant identifier (ordinal factor with 34 levels)
* `trial_number` - trial number within the experiment
* `low_var_color` - the color value of the low-variability item in degrees
* `high_var_color` - the color value of the high-variability item in degrees
* `left_color` - the color value of the left stimulus in degrees
* `right_color` - the color value of the right stimulus in degrees
* `left_sd` - standard deviation of the color distribution for the left stimulus
* `right_sd` - standard deviation of the color distribution for the right stimulus
* `variability_condition` - combined variability condition across the two items
* `similarity_bin` - angular distance between the two item colors in degrees (20°, 45°, or 135°)
* `resp_color` - the color value reported by the participant in degrees
* `rawError` - raw error in the participant's response
* `base_color ` - the actual average color value to be reported
* `realErr` - the actual error in the participant's response
* `errorCorrected` - the error after correction for individual biases (better to use this one for analyses than the raw error)
* `bias_rel_to_high_var` - signed bias, where positive values indicate responses shifted toward the high-variability item
* `is_outlier` - binary indicator of whether the trial was identified as an outlier based on the error magnitude (0 = not an outlier, 1 = outlier)

# Methods

## Participants
`r str_to_sentence(xfun::numbers_to_words(aggr_stats[,.N]))` participants took part in the study. `r str_to_sentence(xfun::numbers_to_words(aggr_stats[,.N]-tomatoes[,lengthu(subject)]))` participants were excluded because of very poor performance (an average absolute error was above the predefined threshold of 60 deg.).

{{< include tomatoes_experience_methods.md >}}

# Results


```{r}
#| fig-cap: "Mean absolute error in color reporting as a function of similarity bin. Error bars represent within-subject 95% confidence intervals."
#| label: fig-color-error
# Create the plot with proper axis labels
plot_pointrange(tomatoes[!is.na(abs_err)&is_outlier==F],
                aes(x = similarity_bin, y = abs_be_c, color = variability_condition), 
                within_subj = TRUE, 
                wid = 'subject', 
                withinvars = c('similarity_bin')
                ) +
  labs(x = "Similarity Bin (degrees)", 
       y = "Absolute Error (degrees)") 

# Run ANOVA
abs_error_anova <- ezANOVA(
  data = tomatoes[!is.na(abs_err)&is_outlier==F],
  dv = .(abs_be_c),
  wid = .(subject),
  within = .(similarity_binf),
  detailed = TRUE,
  type = 3
)



```

The repeated-measures ANOVA showed an effect of similarity bin, `r apa(abs_error_anova, "similarity_binf")`, indicating that participants' performance was affected by the similarity between the two item colors.

```{r}
#| fig-cap: "Bias in reported color as a function of similarity bin and variability condition. Positive values indicate bias toward the high-variability item and negative values indicate bias toward the low-variability item. Error bars represent within-subject 95% confidence intervals."
#| label: fig-bias-analysis

# Calculate bias relative to the high-variability item
tomatoes[!is.na(abs_err), 
         bias_rel_to_high_var_raw := sign(angle_diff_360(item_high_var_color, base_color )) * realErr]

tomatoes[,.(item_low_var_color, item_high_var_color, base_color,overall_sample_mean,resp_color,realErr, bias_rel_to_high_var_raw,  bias_rel_to_high_var )]
# Create the plot
p <- plot_pointrange(tomatoes[is_outlier==F & !is.na(abs_err)], 
                aes(x = similarity_bin, color = variability_condition, y = bias_rel_to_high_var), 
                within_subj = TRUE, 
                wid = 'subject', custom_geom_before = geom_hline(yintercept = 0, linetype = 2),
                design = 'b') +
  labs(x = "Similarity Bin (degrees)", 
       y = "Bias Toward High-Variability Item (degrees)",
       color = "Variability Condition")

p

# save_as_pptx(p, fname = 'fig-tomato-bias.pptx', plot_width = 10, plot_height = 5)
# Run ANOVA on bias
bias_anova <- ezANOVA(
  data = tomatoes[is_outlier==F & !is.na(abs_err),],
  dv = .(bias_rel_to_high_var),
  wid = .(subject),
  within = .(similarity_bin, variability_condition),
  type = 3
)
```

We calculated a bias measure to determine whether participants' color reports were systematically pulled toward one of the two items. This bias was computed so that positive values indicate responses shifted toward the high-variability item, while negative values indicate responses shifted toward the low-variability item.

The analysis of bias revealed that the effect of similarity bin was not significant, `r apa(bias_anova, "similarity_bin")`, indicating that similarity alone did not systematically affect the direction of participants' errors. However, the effect of variability condition was significant, `r apa(bias_anova, "variability_condition")`, demonstrating that variability significantly influenced how participants weighted the two items when computing the average. Importantly, we found a significant interaction between similarity bin and variability condition, `r apa(bias_anova, "similarity_bin:variability_condition")`, indicating that the effect of variability on bias was moderated by item similarity. 

```{r}
#| label: explore-interaction
# Compute mean and CI for each group
mean_bias <- get_superb_ci(
  tomatoes[!is.na(abs_err) & is_outlier==F],   
  wid = c('subject'),  
  within = c('variability_condition', 'similarity_binf'), 
  value_var = 'bias_rel_to_high_var'
)
setDT(mean_bias)

# Add formatted string representation for mean and CI to each row
mean_bias[, mean_ci_str := sprintf("_M_ = %.2f [%.2f, %.2f]", 
                                  center, 
                                  center + lowerwidth, 
                                  center + upperwidth)]

# Split by similarity bin and perform separate ANOVAs
similarity_bins <- unique(tomatoes$similarity_binf)
split_results <- lapply(split(tomatoes[!is.na(abs_err) & is_outlier==F], by = 'similarity_bin'), function(dt) {
  ezANOVA(
    data = dt,
    dv = .(bias_rel_to_high_var),
    wid = .(subject),
    within = .(variability_condition),
    type = 3
  )
})
names(split_results)<-paste0('c',names(split_results))
```

Separate ANOVAs for each similarity bin showed that the effect of variability condition was significant only when items were relatively similar (`r apa(split_results$c20, "variability_condition")` for 20 deg. similarity and `r apa(split_results$c45, "variability_condition")` for 45 deg.) but not when they were very different (`r apa(split_results$c135, "variability_condition")`). With both 20 and 40 similarity, responses were closer to the more variable item, as @fig-bias-analysis shows.

