---
title: The tomato experience
knitr:
  opts_chunk: 
    echo: false
    message: false
    warning: false
    comment: NA
    cache: false
    fig.width: 7
    fig.asp: 0.5
    fig.path: "figures_tomatoes/"
    dpi: 320
    out.width: 100%
    external: false
    res: 320
    results: "hide"
    split: false 
    tidy: false
    dev: svg
editor_options: 
  chunk_output_type: console
format: 
  html: 
    embed-resources: true
params:
  rebuild_data: false
---


```{r setup}
library(httr)
library(apastats2)
library(circhelp)
library(data.table)
library(jsonlite)
library(stringr)
library(knitr)
library(Hmisc)
library(ez)
library(emmeans)
library(superb)

rebuild_data <- exists("params", inherits = FALSE) &&
  is.list(params) &&
  isTRUE(params$rebuild_data)

default_font = 'Gill Sans Nova Light'

default_font_size <- 10
default_line_size <- 1/.pt/3.82*4 
default_font_size_mm <- default_font_size/ggplot2:::.pt

default_point_size <- 4*default_line_size

update_geom_defaults("line", list(linewidth = default_line_size))
update_geom_defaults("text", list(size = default_font_size_mm, family = default_font))
update_geom_defaults("segment", list(linewidth = default_line_size))
update_geom_defaults("pointrange", list(linewidth = default_line_size))
update_geom_defaults("vline", list(linewidth = default_line_size, colour = '#AFABAB'))
update_geom_defaults("hline", list(linewidth = default_line_size, colour = '#AFABAB'))
update_geom_defaults('function', list(linewidth = default_line_size))

default_theme<-theme_light(base_size = default_font_size, base_line_size = default_line_size, base_family = default_font)+theme(
  axis.line=element_line(linewidth = I(0.5)), 
  axis.ticks= element_line(linewidth = I(0.25), colour = 'gray'),
  axis.line.x=element_line(),
  axis.line.y=element_line(),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  legend.title=element_text(size=rel(1)), 
  strip.text=element_text(size=rel(1), colour = 'black'), 
  axis.text=element_text(size=rel(0.9), colour = '#4e4e4e'), 
  axis.title=element_text(size=rel(1), colour = '#4e4e4e'), 
  panel.border= element_blank(),
  strip.background = element_blank(),
  legend.position	='right', 
  legend.key.height = unit(0.7, 'lines'),
  plot.title=element_text(size=rel(1), hjust = 0.5),
  plot.tag.position = c(0,1),
  plot.background = element_blank(),
  text=element_text(size=default_font_size), 
  legend.text=element_text(size=rel(1)), 
  axis.line.x.bottom = element_blank(), 
  axis.line.y.left = element_blank(),
  axis.line.x.top = element_blank(),
  axis.line.y.right = element_blank())
theme_set(default_theme)

default_colors <-c('#3498db','#009f06','#FF7F00', '#7A378B','#f72034')
scale_colour_discrete <- function(...) scale_colour_manual(values=default_colors, ...)
scale_fill_discrete <- function(...) scale_fill_manual(values=default_colors, ...)
scale_shape_discrete <- function(...) scale_shape_manual(values = c(16,21,17, 22), ...)

```

```{r}
#| label: data-download
#| eval: !expr rebuild_data
headers = c(
  `Authorization` = sprintf("Bearer %s",Sys.getenv('JATOS_TOKEN')),
  `Content-Type` = 'text/plain'
)

jatos_url <- 'https://jatos.mindprobe.eu/jatos/api/v1/'
res <- httr::GET(url = paste0(jatos_url, "admin/token"), httr::add_headers(.headers=headers))

randstring <- function(n = 5000) {
  a <- do.call(paste0, replicate(5, sample(LETTERS, n, TRUE), FALSE))
  paste0(a, sprintf("%04d", sample(9999, n, TRUE)), sample(LETTERS, n, TRUE))
}

temp_dir <- tempdir(check = T)
temp_dir <- gsub('\\\\', '/', temp_dir)
temp_dir <- paste(temp_dir, randstring(1), sep = '/')
res_file <- paste0(tempfile(),'.zip')

study_uid <- Sys.getenv('JATOS_TOMATO_STUDY_2026_UID')

req_url <-  sprintf("%sresults/metadata?studyUuid=%s", jatos_url, study_uid)
system(sprintf('curl  -H "Authorization: %s" %s --output "%s" ', headers[['Authorization']], req_url, res_file))

metadata_global <- jsonlite::fromJSON(res_file)$data$studyResults[[1]]
setDT(metadata_global)
unlink(paste0(temp_dir, "/*"))
finished_study <- metadata_global[studyState=='FINISHED',c('id','startDate','endDate','duration')]
finished_study <- finished_study[as.POSIXct(startDate/1000, origin="1970-01-01")>Sys.getenv("JATOS_START_DATE")]
url <-  sprintf("%sresults?studyUuid=%s", jatos_url, study_uid)
system(sprintf('curl -i -H "Authorization: %s" %s --output "%s"', headers[['Authorization']], url, res_file))
unzip(res_file, exdir=temp_dir)  # unzip your file

metadata <- fromJSON(list.files(temp_dir, '*.json', full.names = T, recursive = T))$data$studyResults[[1]]

finished_study_ids <- metadata[,c('id', "endDate", "startDate")]
setDT(finished_study_ids)
finished_study_ids[,id:=as.character(id)]
finished_study_ids[,endDate:=as.POSIXct(endDate/1000, origin="1970-01-01")]
finished_study_ids[,startDate:=as.POSIXct(startDate/1000, origin="1970-01-01")]
finished_study_ids <- finished_study_ids[startDate>Sys.getenv("JATOS_START_DATE")]

comp_res_meta <- rbindlist(lapply(1:length(metadata$componentResults), \(x) {
  dt <- as.data.table(metadata$componentResults[[x]])
  dt
}), fill = T)
flist <- list.files(temp_dir, '*.txt', full.names = T, recursive = T)
flist <- flist[str_extract(flist,'study_result_(\\d+)', group = 1)%in%finished_study_ids$id]

cur_data <- rbindlist(lapply(c(flist), \(x){
  print(x)
  comp_res_id <- str_extract(x, '(?<=comp-result_)\\d{6}')
  data <- read_json(x)
  if (length(data)<40){
    return(data.frame())
  }
  browser_check <-data[sapply(data, \(x) x$trial_type=='browser-check')][[1]]

  setDT(browser_check)
  
  data <- data[sapply(data, \(x) x$trial_type=='psychophysics')]
  data <- rbindlist(data, fill = T)
  if ('score' %nin% colnames(data)){
    return(data.frame())
  }

  km <- kmeans(as.matrix(data[!is.na(click_x),.(click_x,click_y)]), 3, nstart = 50)
  sorted_indices <- order(km$centers[, 1])  # Assuming x-coordinate is in the first column
  sorted_indices[2:3] <- rev(sorted_indices[2:3])
      
  # Reassign cluster labels based on sorted order
  new_clusters <- match(km$cluster, sorted_indices)
  data[!is.na(click_x), cue_n_alt:=new_clusters]
  # data[, cue_pos:=ifelse(cue_n==1, 'left', ifelse(cue_n==2, 'right', 'avg'))]
    
  if ('cue_n' %nin% names(data)){
    data[,cue_n:=cue_n_alt]
  } 
  data[,score:=as.numeric(score)]
  data[, c('left_sd', 'right_sd', 'left_sample_mean_check', 'right_sample_mean_check', 'overall_sample_mean_check') := {
    left_vals <- as.numeric(str_split(left_colors, ',', simplify = TRUE))
    right_vals <- as.numeric(str_split(right_colors, ',', simplify = TRUE))
    .(
      circ_sd_360(left_vals),
      circ_sd_360(right_vals),
      circ_mean_360(left_vals),
      circ_mean_360(right_vals),
      circ_mean_360(c(left_vals, right_vals))
    )
  }, by = .(left_colors, right_colors)]
  data[,left_colors:=fifelse(is.na(left_colors), shift(left_colors), left_colors)]
  data[,right_colors:=fifelse(is.na(right_colors), shift(right_colors), right_colors)]
  setnafill(data, type = "locf",cols = c('target_color','distractor_color','left_color','right_color','left_population_mean','right_population_mean', 'left_correlation', 'right_correlation','base_ori','overall_sample_mean','similarity', 'similarity_bin', 'color_bin','left_sd','right_sd', 'left_sample_mean_check', 'right_sample_mean_check', 'overall_sample_mean_check','cue_n'))
  setnafill(data, type = "nocb",cols = c('score'))
  setnames(data,c('base_ori','left_color','right_color'),c('base_color','left_sample_mean','right_sample_mean'))
  #
  if (browser_check[1, .(trial_type)]=='browser-check'){
    data<-cbind(data,browser_check[,.(width, height, fullscreen)])
  }
  data<-data[trial_type=='psychophysics'&seq_type=='response']
  # data[, error:=angle_diff_360(resp_color, target_color)]
  data[,subject_id:=str_extract(x, 'study_result_(\\d+)', group = 1)]
  data[,result_id:=str_extract(x, 'comp-result_(\\d+)', group = 1)]
  data <- merge(data, finished_study_ids, by.x = 'subject_id', by.y = 'id', all.x = T)
  
  data
  
}), fill = T)
```


```{r}
#| label: corr_matrices_plots
#| eval: false
cur_data[,cor_side:=fifelse(left_correlation>right_correlation, 'left', 'right')]
cur_data[, base_color := atan2(
  sin(target_color / 180 * pi) + sin(distractor_color / 180 * pi),
  cos(target_color / 180 * pi) + cos(distractor_color / 180 * pi)
) / pi * 180]

cur_data[, base_color := angle_diff_360(base_color,0)]
# dt <- cur_data[left_correlation==0.8&right_correlation==0]
lapply(split(cur_data, by = c('left_correlation', 'right_correlation'), drop = T), \(dt){
  trial_colors <- as.matrix(dt[,tstrsplit(paste(left_colors, right_colors, sep = ','), ',', type.convert = as.numeric)])

  trial_colors[, 1:64] <- sweep(
    trial_colors[, 1:64, drop = FALSE],
    1,
    dt$left_population_mean,
    FUN = angle_diff_360
  )

  trial_colors[, 65:128] <- sweep(
    trial_colors[, 65:128, drop = FALSE],
    1,
    dt$right_population_mean,
    FUN = angle_diff_360
  )

  means_left <- rowMeans(trial_colors[,1:64])
  means_right <- rowMeans(trial_colors[,65:128])
  cor_mat <- cor(trial_colors)
  print(sprintf('Left correlation: %.2f, Right correlation: %.2f', dt$left_correlation[1], dt$right_correlation[1]))
  print(sprintf('Mean correlation of left item colors: %.2f', mean(cor_mat[1:64,1:64][lower.tri(cor_mat[1:64,1:64])])))
  print(sprintf('Mean correlation of right item colors: %.2f', mean(cor_mat[65:128,65:128][lower.tri(cor_mat[65:128,65:128])])))
  print(sprintf('Correlation between left and right item colors: %.2f', mean(cor_mat[1:64,65:128][lower.tri(cor_mat[1:64,65:128])])))
  
  GGally::ggcorr(data  = NULL, cor_matrix = cor_mat)+labs(title= sprintf('Left correlation: %.2f, Right correlation: %.2f', dt$left_correlation[1], dt$right_correlation[1]))
})
```
```{r}
#| label: data-preprocessing
#| eval: !expr rebuild_data
tomatoes <- cur_data
tomatoes[,expName:='tomato_experience_2026']
tomatoes[,.(left_population_mean, right_population_mean, base_color,
            target_color, distractor_color, left_sample_mean, right_sample_mean)]
tomatoes[,error_to_pop_mean:=angle_diff_360(resp_color, base_color )]
tomatoes[left_correlation!=right_correlation, c('error_to_uncorrelated','error_to_correlated') := .(
  angle_diff_360(resp_color, fifelse(left_correlation<right_correlation,  left_sample_mean, right_sample_mean)), 
  angle_diff_360(resp_color, fifelse(left_correlation>right_correlation,  left_sample_mean, right_sample_mean))
  )]

# correlation level defines variability level
# (correlated = low variability, uncorrelated = high variability).
tomatoes[,item_low_var_color:=ifelse(
  left_correlation>right_correlation, left_sample_mean,
  right_sample_mean
)]
tomatoes[,item_high_var_color:=ifelse(
  left_correlation>right_correlation, right_sample_mean,
  left_sample_mean
)]

tomatoes[,abs_err:=abs(error)]
tomatoes <- unique(tomatoes)
tomatoes[!is.na(error),n_by_subj:=.N, by = subject_id]
tomatoes[!is.na(error),n_by_block:=.N, by = .(result_id, subject_id)]

tomatoes[,color_dist:=angle_diff_360(item_low_var_color, item_high_var_color)]
tomatoes[,abs_color_dist:=abs(color_dist)]
tomatoes[,similarity_binf:=factor(similarity_bin)]

tomatoes<-tomatoes[n_by_subj>250&block>0]

aggr_stats <- tomatoes[,.(totalN = .N, non_missing=n_by_subj[1], 
                          circ_sd_err_to_pop_mean = circ_sd_360(error_to_pop_mean, na.rm = T), 
                          circ_sd_err_to_sample_mean = circ_sd_360(error, na.rm = T),
                          circ_sd_err_to_uncorrelated = circ_sd_360(error_to_uncorrelated, na.rm = T),
                          circ_sd_err_to_correlated = circ_sd_360(error_to_correlated, na.rm = T),
                          n_by_subj[1], missed_responses = sum(is.na(error))), keyby = .(subject_id)]
fwrite(tomatoes, file = "tomatoes_raw_2026.csv")                
```

```{r}
#| label: data-quality-checks
#| results: hide
#| eval: false
tomatoes <- fread('tomatoes_raw_2026.csv')
tomatoes[,.(.N, circ_sd = circ_sd_360(error, na.rm = T), mean_abs_err= mean_nn(abs_err)), keyby = .(startDate, endDate,  subject_id)]
ggplot(tomatoes[!is.na(error)], aes(x = error))+
  geom_density()+
  geom_rug()
  labs(x = 'Error (degrees)', y = 'Density')+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

```{r}
#| label: bias-correction-outlier-removal
#| eval: !expr rebuild_data
#outliers 
tomatoes <- tomatoes[subject_id%nin%tomatoes[,.(.N, circ_sd_360(error, na.rm = T)), by = subject_id][V2>60|N<200, subject_id]]
tomatoes[,subject:=paste0('S',as.numeric(as.factor(subject_id)))]
tomatoes[,subject:=naturalsort::naturalfactor(subject)]

tomatoes[,.(.N, circ_sd_360(error, na.rm = T)), by = subject]
db_res <- lapply(split(tomatoes[!is.na(error),], by = c('subject_id')), \(dt){
  dt[,c('be_c','is_outlier','rc_pred','rc_group','shifted_color_dist'):=circhelp::remove_cardinal_biases(error, overall_sample_mean , space = '360', plots = 'hide')[,c('be_c','is_outlier','pred','which_bin','shifted_x')], by = .(subject)]
  dt
})

tomatoes <-rbindlist(db_res)
# tomatoes <- tomatoes[!is.na(abs_err)&is_outlier==F]
```

```{r}
#| label: creating-output-dataset
#| eval: !expr rebuild_data
tomatoes[is_outlier==F&abs_err>90, is_outlier:=T]
corr_midpoint <- mean(range(c(tomatoes$left_correlation, tomatoes$right_correlation), na.rm = TRUE))
tomatoes[,left_variability := fifelse(left_correlation <= corr_midpoint, "high", "low")]
tomatoes[,right_variability := fifelse(right_correlation <= corr_midpoint, "high", "low")]
tomatoes[,noise_equal:=factor(left_variability == right_variability, levels = c(FALSE, TRUE), labels = c('Unequal noise', 'Equal noise'))]
tomatoes[,variability_condition:=factor(paste0(left_variability, " - ", right_variability),
                                        levels = c("low - low", "high - low", "low - high", "high - high"))]
tomatoes[,similarity_binf:=factor(similarity_bin)]

set.seed(20260425)
tomatoes[, item_a_is_left := sample(c(TRUE, FALSE), .N, replace = TRUE)]
tomatoes[, item_a_color := fifelse(item_a_is_left, left_sample_mean, right_sample_mean)]
tomatoes[, item_b_color := fifelse(item_a_is_left, right_sample_mean, left_sample_mean)]
tomatoes[, item_a_variability := fifelse(item_a_is_left, left_variability, right_variability)]
tomatoes[, item_b_variability := fifelse(item_a_is_left, right_variability, left_variability)]
tomatoes[, variability_condition_ab := factor(
  paste0(item_a_variability, " - ", item_b_variability),
  levels = c("low - low", "high - low", "low - high", "high - high")
)]
tomatoes[!is.na(abs_err),
         bias_rel_to_item_a := sign(angle_diff_360(item_a_color, base_color)) * angle_diff_360(resp_color, base_color )]

aggr_stats <- tomatoes[,.(.N, circ_sd = circ_sd_360(error, na.rm = T),circ_sd_correct = circ_sd_360(be_c, na.rm = T), n_by_subj=n_by_subj[1], n_nonmissing=sum(!is.na(error)), share_outliers= mean_nn(is_outlier), mean_abs_err= mean_nn(abs_err), bias_rel_to_item_a = circ_mean_360(bias_rel_to_item_a, na.rm = T)), keyby = .(subject)]#[,mean_nn(share_outliers)]

tomatoes[,abs_be_c:=abs(be_c)]
tomatoes[,avg_abs_err:=mean_nn(abs(be_c)), by = subject]

tomatoes <- drop_empty_cols(tomatoes)

tomatoes[,c('startDate','endDate','result_id','subject_id','avg_frame_time'):=NULL]

if (lengthu(tomatoes$subject)>1){
  tomatoes <- tomatoes[, lapply(.SD, function(col) if (!all(na.omit(col == col[!is.na(col)][1]))) col else NULL)]
}
tomatoes_out <- tomatoes[order(subject, trial_number),.(subject, trial_number, low_var_color = item_low_var_color, high_var_color = item_high_var_color, left_color = left_sample_mean, right_color = right_sample_mean, left_sd,  right_sd, variability_condition, variability_condition_ab, item_a_is_left, item_a_color, item_b_color, similarity_bin, resp_color,  error_to_sample_mean = error, base_color , error_to_pop_mean, errorCorrected = be_c, bias_rel_to_item_a, is_outlier)]
saveRDS(tomatoes, file = 'tomatoes_preprocessed_2026.rds')
fwrite(tomatoes_out, file = 'tomatoes_2026.csv')
fwrite(aggr_stats, file = 'tomatoes_aggr_stats.csv')
```

# Instructions

Read the report below and use it as a basis for your presentation. You can also download the preprocessed data from `tomatoes_2026.csv` and use your favorite software (JASP, jamovi, R, SPSS) to analyze it to get more insight into the study results. The data file has the following columns:

* `subject` - participant identifier (ordinal factor with 34 levels)
* `trial_number` - trial number within the experiment
* `low_var_color` - the color value of the low-variability item in degrees
* `high_var_color` - the color value of the high-variability item in degrees
* `left_color` - the color value of the left stimulus in degrees
* `right_color` - the color value of the right stimulus in degrees
* `left_sd` - standard deviation of the color distribution for the left stimulus
* `right_sd` - standard deviation of the color distribution for the right stimulus
* `variability_condition` - combined variability condition across the two items
* `variability_condition_ab` - combined variability condition after randomly assigning the two items to A and B on each trial
* `item_a_is_left` - logical indicator of whether item A is the left item on that trial
* `item_a_color` - the color value of item A in degrees
* `item_b_color` - the color value of item B in degrees
* `similarity_bin` - angular distance between the two item colors in degrees (20°, 45°, or 135°)
* `resp_color` - the color value reported by the participant in degrees
* `error_to_sample_mean` - raw error in the participant's response
* `base_color ` - the actual average color value to be reported
* `error_to_pop_mean` - the actual error in the participant's response
* `errorCorrected` - the error after correction for individual biases (better to use this one for analyses than the raw error)
* `bias_rel_to_item_a` - signed bias, where positive values indicate responses shifted toward item A and negative values indicate responses shifted toward item B
* `is_outlier` - binary indicator of whether the trial was identified as an outlier based on the error magnitude (0 = not an outlier, 1 = outlier)

```{r}
#| label: read-preprocessed-data
aggr_stats <- fread('tomatoes_aggr_stats.csv')
tomatoes <- readRDS('tomatoes_preprocessed_2026.rds')
```

# Methods

## Participants
`r str_to_sentence(xfun::numbers_to_words(aggr_stats[,.N]))` participants took part in the study. `r str_to_sentence(xfun::numbers_to_words(aggr_stats[,.N]-tomatoes[,lengthu(subject)]))` participants were excluded because of very poor performance (an average absolute error was above the predefined threshold of 60 deg.).

{{< include tomatoes_experience_methods_V26.md >}}

# Results


```{r}
#| fig-cap: "Mean absolute error in color reporting as a function of similarity bin. Error bars represent within-subject 95% confidence intervals."
#| label: fig-color-error
# Create the plot with proper axis labels
plot_pointrange(tomatoes[!is.na(abs_err)&is_outlier==F],
                aes(x = similarity_bin, y = abs_err, color = variability_condition), 
                within_subj = TRUE, 
                wid = 'subject', 
                withinvars = c('similarity_bin')
                ) +
  labs(x = "Similarity Bin (degrees)", 
       y = "Absolute Error (degrees)",
       color = "Variability Condition") 

# Run ANOVA
abs_error_anova <- afex::aov_ez(
  data = tomatoes[!is.na(abs_err)&is_outlier==F],
  dv = 'abs_be_c',
  id = 'subject',
  within = c('similarity_bin', 'variability_condition'),
  type = 3, fun_aggregate = mean
)

means_by_var <- get_superb_ci(
  tomatoes[!is.na(abs_err) & is_outlier==F],   
  wid = c('subject'),  
  within = c('variability_condition'), 
  value_var = 'abs_be_c'
)
setDT(means_by_var)
```
```{r}
#| label: posthoc-tests-absolute-error
posthoc_tests <- summary(contrast(emmeans(abs_error_anova,  ~  variability_condition), 'pairwise'))
kable(posthoc_tests)
```

The repeated-measures ANOVA showed an effect of similarity bin, `r apa(abs_error_anova, "similarity_bin")`, indicating that participants' performance was affected by the similarity between the two item colors. Variability condition also had a significant effect on absolute error, `r apa(abs_error_anova, "variability_condition")`, as expected, with low-low condition (`r means_by_var[variability_condition=='low - low', descr]`; here and later 95% confidence intervals are reported in brackets after the means) being below the other three (`r paste_and(means_by_var[variability_condition!='low - low', descr])` for `r paste_and(means_by_var[variability_condition!='low - low', variability_condition])`, respectively). 

```{r}
#| fig-cap: "Bias toward the higher-variability item as a function of similarity bin, computed on unequal-variance trials only. Positive values indicate responses shifted toward the higher-variability item. Error bars represent within-subject 95% confidence intervals."
#| label: fig-bias-analysis

# Collapse across high-low and low-high trials by coding bias toward the higher-variability item.
cols <- c('overall_sample_mean', 'resp_color', 'item_high_var_color')
tomatoes[, (cols) := lapply(.SD, angle_diff_360, 0), .SDcols = cols]
tomatoes[!is.na(abs_err) & left_variability != right_variability,
         bias_rel_to_high_var_raw := sign(angle_diff_360(item_high_var_color, overall_sample_mean)) * be_c]

uneq_bias <- tomatoes[!is.na(abs_err) & is_outlier==F & left_variability != right_variability]

p <- plot_pointrange(uneq_bias,
                aes(x = similarity_bin, y = bias_rel_to_high_var_raw), 
                within_subj = TRUE, 
                wid = 'subject', custom_geom_before = geom_hline(yintercept = 0, linetype = 2),
                design = 'w') +
  labs(x = "Similarity Bin (degrees)", 
       y = "Bias Toward Higher-Variability Item (degrees)")

p

bias_anova <- afex::aov_ez(
  data = uneq_bias,
  dv = 'bias_rel_to_high_var_raw',
  id = 'subject',
  within = c('similarity_bin'),
  type = 3,
  fun_aggregate = mean
)

subject_bias <- uneq_bias[, .(bias_rel_to_high_var_raw = mean(bias_rel_to_high_var_raw, na.rm = TRUE)), by = .(subject, similarity_bin)]
mean_bias <- get_superb_ci(
  subject_bias,
  wid = c('subject'),
  within = c('similarity_bin'),
  value_var = 'bias_rel_to_high_var_raw'
)
setDT(mean_bias)

bias_ttests <- subject_bias[, .(apa = apa(t.test(bias_rel_to_high_var_raw, mu = 0))), by = similarity_bin]

```

We calculated a bias measure to determine whether participants' reports were systematically pulled toward the higher-variability item. For this analysis, we included only unequal-variance trials and coded the signed error so that positive values indicate responses shifted toward the higher-variability item.

The one-way repeated-measures ANOVA showed no effect of similarity bin, `r apa(bias_anova, "similarity_bin")`. One-sample t-tests showed a significant positive bias only at 20 deg. similarity (`r mean_bias[similarity_bin == 20, descr]`, `r bias_ttests[similarity_bin == 20, apa]`). The corresponding tests were not significant at 45 deg. (`r mean_bias[similarity_bin == 45, descr]`, `r bias_ttests[similarity_bin == 45, apa]`) or at 135 deg. (`r mean_bias[similarity_bin == 135, descr]`, `r bias_ttests[similarity_bin == 135, apa]`).

```{r}
#| fig-cap: "Individual differences in mean absolute error and mean bias toward the higher-variability item. Each point is one observer; bias is averaged across unequal-variance trials. The line shows the least-squares fit."
#| label: fig-bias-accuracy-individual-differences

subject_bias_overall <- uneq_bias[, .(mean_bias = mean(bias_rel_to_high_var_raw, na.rm = TRUE)), by = subject]
subject_abs_err <- tomatoes[!is.na(abs_be_c) & is_outlier==F, .(mean_abs_err = mean(abs_be_c, na.rm = TRUE)), by = subject]
individual_differences <- merge(subject_bias_overall, subject_abs_err, by = 'subject')
indiv_diff_cor <- cor.test(individual_differences$mean_abs_err, individual_differences$mean_bias)

ggplot(individual_differences, aes(x = mean_abs_err, y = mean_bias)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point() +
  geom_smooth(method = MASS::rlm, se = FALSE) +
  labs(
    x = "Mean Absolute Error (degrees)",
    y = "Mean Bias Toward Higher-Variability Item (degrees)"
  )
```

At the individual-differences level, observers with larger mean absolute error tended to show less positive bias toward the higher-variability item, `r apa(indiv_diff_cor)`. So in these data, less accurate observers were less biased toward the higher-variability item.

