---
title: Cueing Temporal Attention
knitr:
  opts_chunk: 
    echo: false
    message: false
    warning: false
    comment: NA
    cache: false
    fig.width: 7
    fig.asp: 0.5
    fig.path: "figures/"
    dpi: 320
    external: false
    res: 320
    results: "hide"
    out.width: 100%
    split: false 
    tidy: false
    dev: svg
editor_options: 
  chunk_output_type: console
format: 
  html: 
    embed-resources: true
---


```{r setup}
library(httr)
library(apastats)
library(data.table)
library(jsonlite)
library(stringr)
library(knitr)
library(patchwork)
library(Hmisc)
library(superb)
library(ez)
library(circhelp)

default_font = 'Gill Sans Nova Light'

default_font_size <- 10
default_line_size <- 1/.pt/3.82*4 
default_font_size_mm <- default_font_size/ggplot2:::.pt

default_point_size <- 4*default_line_size

update_geom_defaults("line", list(linewidth = default_line_size))
update_geom_defaults("text", list(size = default_font_size_mm, family = default_font))
update_geom_defaults("segment", list(linewidth = default_line_size))
update_geom_defaults("pointrange", list(linewidth = default_line_size))
update_geom_defaults("vline", list(linewidth = default_line_size, colour = '#AFABAB'))
update_geom_defaults("hline", list(linewidth = default_line_size, colour = '#AFABAB'))
update_geom_defaults('function', list(linewidth = default_line_size))

default_theme<-theme_light(base_size = default_font_size, base_line_size = default_line_size, base_family = default_font)+theme(
  axis.line=element_line(linewidth = I(0.5)), 
  axis.ticks= element_line(linewidth = I(0.25), colour = 'gray'),
  axis.line.x=element_line(),
  axis.line.y=element_line(),
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(),
  legend.title=element_text(size=rel(1)), 
  strip.text=element_text(size=rel(1), colour = 'black'), 
  axis.text=element_text(size=rel(0.9), colour = '#4e4e4e'), 
  axis.title=element_text(size=rel(1), colour = '#4e4e4e'), 
  panel.border= element_blank(),
  strip.background = element_blank(),
  legend.position	='right', 
  legend.key.height = unit(0.7, 'lines'),
  plot.title=element_text(size=rel(1), hjust = 0.5),
  plot.tag.position = c(0,1),
  plot.background = element_blank(),
  text=element_text(size=default_font_size), 
  legend.text=element_text(size=rel(1)), 
  axis.line.x.bottom = element_blank(), 
  axis.line.y.left = element_blank(),
  axis.line.x.top = element_blank(),
  axis.line.y.right = element_blank())
theme_set(default_theme)

default_colors <-c('#3498db','#009f06','#FF7F00', '#7A378B','#f72034')
scale_colour_discrete <- function(...) scale_colour_manual(values=default_colors, ...)
scale_fill_discrete <- function(...) scale_fill_manual(values=default_colors, ...)
scale_shape_discrete <- function(...) scale_shape_manual(values = c(16,21,17, 22), ...)


gg_crosshair <- function(xint = 0, yint = 0, lty = 2) {
  if (!is.null(xint)) gv = geom_vline(xintercept = xint, linetype = lty)
  else gv = NULL
  if (!is.null(yint)) gh = geom_hline(yintercept = yint, linetype = lty)
  else gh = NULL
  
  list(gv, gh)
}
headers = c(
  `Authorization` = sprintf("Bearer %s", Sys.getenv("JATOS_TOKEN"))
)

jatos_url <- 'https://jatos.mindprobe.eu/jatos/api/v1/'

randstring <- function(n = 5000) {
  a <- do.call(paste0, replicate(5, sample(LETTERS, n, TRUE), FALSE))
  paste0(a, sprintf("%04d", sample(9999, n, TRUE)), sample(LETTERS, n, TRUE))
}

data_list <- list()
study_uid <- Sys.getenv('JATOS_TEMP_ATT_STUDY_UID')

temp_dir <- tempdir(check = T)
temp_dir <- gsub('\\\\', '/', temp_dir)
temp_dir <- paste(temp_dir, randstring(1), sep = '/')
res_file <- paste0(tempfile(),'.zip')

url <-  sprintf("%sresults?studyUuid=%s", jatos_url, study_uid)
system(sprintf('curl -i -H "Authorization: %s" %s --output "%s"', headers[['Authorization']], url, res_file))
unzip(res_file, exdir=temp_dir)  # unzip your file


metadata <- jsonlite::fromJSON(list.files(temp_dir, '*.json', full.names = T, recursive = T))$data$studyResults[[1]]

finished_study_ids <- metadata[,c('id', "endDate", "startDate")]
setDT(finished_study_ids)
finished_study_ids[,id:=as.character(id)]
finished_study_ids[,endDate:=as.POSIXct(endDate/1000, origin="1970-01-01")]
finished_study_ids[,startDate:=as.POSIXct(startDate/1000, origin="1970-01-01")]
finished_study_ids <- finished_study_ids[startDate>Sys.getenv("JATOS_START_DATE")]

comp_res_meta <- rbindlist(lapply(1:length(metadata$componentResults), \(x) {
  dt <- as.data.table(metadata$componentResults[[x]])
  dt
}), fill = T)

flist <- list.files(temp_dir, '*.txt', full.names = T, recursive = T)
flist <- flist[str_extract(flist,'study_result_(\\d+)', group = 1)%in%finished_study_ids$id]

data <- suppressWarnings(rbindlist(lapply(c(flist), \(x){
  print(x)
  comp_res_id <- str_extract(x, '(?<=comp-result_)\\d{6}')
  data <- read_json(x)
  if (length(data)<10){
    return(data.frame())
  }
  data <- data[sapply(data, \(x) x$trial_type=='psychophysics')]
  
  data <- rbindlist(data, fill = T)

  data[,subject_id:=str_extract(x, 'study_result_(\\d+)', group = 1)]
  data[,result_id:=str_extract(x, 'comp-result_(\\d+)', group = 1)]
  data <- merge(data, finished_study_ids, by.x = 'subject_id', by.y = 'id')
  
  data
  
}), fill = T))

unlink(paste0(temp_dir, "/*"))
setnafill(data, type = "nocb", cols = c('start_ori','resp_ori','error','trial_number'))

data[,participant:=as.numeric(factor(subject_id, levels = sample(unique(data$subject_id)), labels = 1:length(unique(data$subject_id))))]
data[,c('startDate','endDate','result_id','subject_id','avg_frame_time'):=NULL]
data <- data[seq_type=='stimuli']


data[,abs_err:=abs(error)]
datam <- melt(data[,.(participant, block, trial_index, t1_ori, t2_ori, t3_ori, real_target_position)], measure.vars = patterns('_ori'))
datam[,item_pos:=as.numeric(str_extract( variable,'\\d+'))]
datam <- datam[order(participant, block, trial_index, item_pos)]
datam[,c('diff_to_next','diff_to_prev'):=list(angle_diff_180(value,shift(value, -1)),angle_diff_180(value,shift(value, 1))), by = .(participant, block, trial_index)]
data<-merge(data, datam[real_target_position==item_pos, .(participant, block, trial_index, target_ori = value, diff_to_next, diff_to_prev)], by = c('participant', 'block', 'trial_index'))
data <- drop.empty.cols(data)
data[,N_by_participant:=.N, by = participant]
data <- data[N_by_participant>100]
data <- data[, lapply(.SD, function(col) if (!all(na.omit(col == col[!is.na(col)][1]))) col else NULL)]

fwrite(data[,.(participant, block, trial_number, t1_ori, t2_ori, t3_ori, target_position, cue_type, real_target_position, start_ori, resp_ori, error, abs_err, target_ori, diff_to_next, diff_to_prev)], file = 'temporal_attention.csv')
```

# Instructions

Read the report below and use it as a basis for your presentation. You can also download the preprocessed data from `temporal_attention.csv` and use your favorite software (JASP, jamovi, R, SPSS) to analyze it to get more insight into the study results. The data file has the following columns:

* `participant` - participant identifier (integer: 1-39)
* `block` - experimental block (0-1)
* `trial_number` - trial number within the experiment
* `t1_ori` - orientation of the first Gabor patch (degrees)
* `t2_ori` - orientation of the second Gabor patch (degrees)
* `t3_ori` - orientation of the third Gabor patch (degrees)
* `target_position` - initial cue position (1, 2, or 3), indicating which position was cued before stimulus presentation
* `cue_type` - validity of the cue ("valid", "invalid", or "neutral")
* `real_target_position` - position that was actually tested (1, 2, or 3), indicating the post-cue
* `start_ori` - starting orientation of the response line (degrees)
* `resp_ori` - orientation reported by the participant (degrees)
* `error` - signed error between response and target orientation (degrees)
* `abs_err` - absolute error between response and target orientation (degrees)
* `target_ori` - the actual orientation of the target Gabor patch (degrees)
* `diff_to_next` - angular difference between the target orientation and the orientation of the next item in sequence (degrees)
* `diff_to_prev` - angular difference between the target orientation and the orientation of the previous item in sequence (degrees)


{{< include temporal_attention_methods.md >}}
### Participants
`r stringr::str_to_sentence(apastats:::numbers2words(data[,lengthu(participant)]))` participants took part in this study. Participation required a web browser with a minimum screen resolution of 1000 × 600 pixels. Demographic information was not collected.

## Results

### Effect of Cue Type on Accuracy

```{r cue-type-effect}
#| label: fig-cue-type-effect
#| fig-cap: "Effect of cue type on orientation judgment accuracy"

data_descr <- data[,.(.N, acc = mean(abs_err)), keyby = .(participant, cue_type)]

# Run ANOVA for cue type effect
cue_type_anova <- ezANOVA(
  data = data,
  dv = .(abs_err),
  wid = .(participant),
  within = .(cue_type),
  detailed = TRUE,
  type = 3
)

# Plot cue type effect using plot.pointrange with direct parameters
plot.pointrange(data, 
                aes(x = cue_type, y = abs_err), 
                within_subj = TRUE,  
                withinvars = "cue_type",
                wid = "participant",
                pointsize = 3,
                linesize = 1)+labs(
                x = "Cue Type", 
                y = "Absolute Error (degrees)")
```

The effect of a pre-cue type on accuracy was analyzed using a within-subject ANOVA with absolute error in orientation judgment as the dependent variable (@fig-cue-type-effect). The results show a clear effect of cue type: `r describe.ezanova(cue_type_anova, 'cue_type')`. On average, participants average absolute error was `r data_descr[,describe.mean.conf(acc), by = .(cue_type)][,paste_and(sprintf('%s with %s cues', V1, cue_type))]` (here and later the confidence intervals are reported in square brackets after the means).


### Effect of Pre-Cue and Post-Cue on Accuracy

```{r cue-positions}
#| label: fig-cue-positions
#| fig-cap: "Effect of initial cue position and post-cue position on accuracy"

# For this analysis, target_position is the initial cue and real_target_position is the post-cue
data[, initial_cue := target_position]

data[, post_cue := real_target_position]
data[, position_label := paste0(initial_cue, "→", post_cue)]
data[, position_label := factor(position_label)]
data[, position_label2:=paste0(initial_cue, "-", post_cue)]
data[,initial_cue:=as.character(initial_cue)]
data[cue_type=='neutral', initial_cue := 'neutral']
# Run ANOVA for position effects (excluding neutral cues)
position_anova <- ezANOVA(
  data = data[cue_type != "neutral"],
  dv = .(abs_err),
  wid = .(participant),
  within = .(initial_cue, post_cue),
  type = 3
)


# Plot the interaction using plot.pointrange
plot.pointrange(data, 
                aes(color = initial_cue, y = abs_err, x = post_cue),
                within_subj = TRUE,
                withinvars = c("initial_cue", "post_cue"),
                wid = "participant",
                pointsize = 3,
                linesize = 1)+labs(
                x = "Post-cue", color = "Pre-cue", 
                y = "Absolute Error (degrees)") 

pre_post <- data[post_cue=='2'&initial_cue!='2'&cue_type != "neutral"]
pre_post_by_subj <- pre_post[,mean.nn(abs_err), by = .(participant, post_cue, initial_cue)]
ttest_descr <- describe.ttest(dcast(pre_post_by_subj, participant~initial_cue, fun=mean)[,t.test(`1`,`3`, paired = T)])

```

The effects of initial cue position and post-cue position were analyzed using a within-subject ANOVA (@fig-cue-positions) after excluding the trials where the first cue was neutral. The results showed:

* a significant main effect of initial cue position: `r describe.ezanova(position_anova, 'initial_cue')`
* a significant main effect of post-cue position: `r describe.ezanova(position_anova, 'post_cue')`
* a significant interaction effect: `r describe.ezanova(position_anova, 'initial_cue:post_cue')`

To examine whether performance drops more when people have to report the item at the position after the one that is cued (i.e., if there is a 'refractory' period for attention), we selected the trials where participants reported the target at position 2 and the pre-cue pointed to the 1st or the 3rd position. We found that after being cued to position 1, participants' reports for the second target suffered more, than when they were initially cued to position 3 (`r ttest_descr`). 

```{r}
#| label: fig-benefit
#| fig-cap: "Cost and Benefit Indices for Temporal Attention"
costs_benefits <- dcast(data[,.(perf = mean.nn(abs_err)), by = .(participant, cue_type, post_cue)], participant+post_cue~cue_type)
costs_benefits[, cost_index := -(neutral-invalid)/(valid-invalid)]
costs_benefits[, benefit_index := -(valid-neutral)/(valid-invalid)]
costs_benefits[post_cue==2,t.test(neutral,invalid, paired  = T)]

costs_benefits <- melt(costs_benefits, id.vars = c('participant', 'post_cue'), measure.vars = c('cost_index', 'benefit_index')) 
costs_benefits[post_cue==1]
costs_benefits[,index_type := ifelse(variable=='cost_index', 'cost', 'benefit')]
cb_tt <- costs_benefits[,describe.ttest(t.test(value)), by = .(variable, post_cue)]
plot.pointrange(costs_benefits, aes(x = post_cue, y = value, color = index_type), 
                  within_subj = TRUE,
                  design = 'w',
                  wid = "participant",
                  custom_geom_before = list(gg_crosshair(yint = 0)),
                  pointsize = 3,
                  linesize = 1) + labs(
                    x = "Reported target", color = "Index Type",
                    y = "Cost/Benefit Index"
                  )


```

Finally, we repeated the cost-benefit analysis introduced by Denison et al. (2017). As @fig-benefit shows, for target 1 and 3, there were reliable costs (`r cb_tt[variable=='cost_index'&post_cue%in%c(1,3), paste_and(V1)]`) along with some benefits that did not reach significance (`r cb_tt[variable=='benefit_index'&post_cue%in%c(1,3), paste_and(V1)]`). No signficant effects were found for target 2 in this type of analysis.

